---
title: "Strawberries-Presentation"
author: "Runqi(Ricky) Zhao"
date: "10/18/2020"
output:
  revealjs::revealjs_presentation:
    theme: solarized
    transition: default
    reveal_options:
      slideNumber: true
---



```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(magrittr)
library(kableExtra)
library(gridExtra)
opts_chunk$set(echo = FALSE, 
               warning = FALSE,
               message = FALSE)
```

# Acquire and read the data
<br>

<font size ="4"> <p align="left">
These data were collected from the USDA database selector: <a href="https://quickstats.nass.usda.gov">https://quickstats.nass.usda.gov</a></p>
<p align="left">
The data were <a href="https://quickstats.nass.usda.gov/results/D416E96E-3D5C-324C-9334-1D38DF88FFF1">stored online</a> and then downloaded as a CSV file.</p>
</font>



## Read data and delete meaningless columns
<br><br>

<font size ="5"> <p align="left">
&ensp; &ensp; &ensp; First look at the columns in the database.
</p></font>
```{r}
# read data
berries <- read_csv("berries.csv", col_names = TRUE)
colnames(berries)
# unique values 
aa <- summarize_all(berries, n_distinct)

bb <- which(aa[1,]==1)

# colnames(berries)[bb]

```

----
<br><br>

<font size ="5"> <p align="left">
&ensp; &ensp; &ensp;The NASS database has columns without any data or with a single repeated Values. We can delete these meaningless columns.
</p></font>

```{r}
# single_values <- berries[1,bb]
# single_values

# meaningful data
berries <- select(berries, -all_of(bb))
aa <- select(aa, -all_of(bb))

# delate state ANSI code
berries <- select(berries, -4)
aa <- select(aa, -4) 

kable_styling(kable(head(berries, n=6)), font_size=12)

```


```{r}
ncol <- dim(berries)[2]
nberry <- length(unique(berries$Commodity))
```

<font size ="5"> <p align="left">
&ensp; &ensp; &ensp;Now we have `r ncol` meaningful columns in this table. 
</p></font>

----

## Strawberries Data
<br>

<font size ="5"> <p align="left">
&ensp; &ensp; &ensp;This table contains `r nberry` berries: blueberries, raspberries, and strawberries. Then working along with strawberries. Only consider the "YEAR" time period.
</p></font>
<font size ="5"> <p align="left">
&ensp; &ensp; &ensp;Since we focus on YEAR period strawberries data, we can delete these two columns from our database.
</p></font>

```{r}
# strawberry data
stb <- filter(berries, (Commodity=="STRAWBERRIES") & (Period=="YEAR"))

stb <- select(stb, -c(Period, Commodity))

kable_styling(kable(head(stb, n=6)), font_size=12)

```

# Clean and organize data
<br>
<font size ="5"> <p align="left">
&ensp; &ensp; &ensp;1. Separate data and delete redundancy.
</p></font>
<font size ="5"> <p align="left">
&ensp; &ensp; &ensp;2. Reorganize the data.
</p></font>

## Separate data and delete redundancy
<br>
<font size ="5"> <p align="left">
&ensp; &ensp; &ensp;Data Item column includes information of berry name, production type, and measurement. Separate them into different columns.
</p></font>

```{r}

# sum(str_detect(stb$`Data Item`, "^STRAWBERRIES")) == length(stb$`Data Item`)
# unique(stb$`Data Item`)

# #two subset
s1 <- str_detect(stb$`Data Item`, "^STRAWBERRIES, ")
s1 <- which(s1 == TRUE)
stb1 <- stb[s1,]
# sum(str_detect(stb1$`Data Item`, "^STRAWBERRIES, ")) == length(stb1$`Data Item`)

s2 <- str_detect(stb$`Data Item`, "^STRAWBERRIES - ")
s2 <- which(s2 == TRUE)
stb2 <- stb[s2,]
# sum(str_detect(stb2$`Data Item`, "^STRAWBERRIES - ")) == length(stb2$`Data Item`)


# unique(stb1$`Data Item`)

stb1 <- separate(stb1, `Data Item`, c("Berry","type", "measure", "other"), sep = ",") 
# unique(stb1$Berry)
# unique(stb1$type)
# unique(stb1$measure)
s11 <- str_detect(stb1$measure, "^ MEASURED")

s12 <- which(s11 == FALSE)
stb12 <- stb1[s12,]
stb12$type <- paste(stb12$type, stb12$measure, sep= ",")
stb12$measure <- stb12$other
# sum(str_detect(stb12$measure, "^ MEASURED")) == length(stb12$measure)

stb12$other <- NA

s11 <- which(s11 == TRUE)
stb11 <- stb1[s11,]

stb1 <- rbind(stb11,stb12)

# unique(stb1$Berry)
# unique(stb1$type)
# unique(stb1$measure)
# unique(stb1$other)


stb2 <- separate(stb2, `Data Item`, c("Berry","type"), sep = "-") 
# unique(stb2$Berry)
# unique(stb2$type)
stb2 <- separate(stb2, `type`, c("type", "measure", "other"), sep = ",") 
# unique(stb2$measure)
# unique(stb2$other)

stb <- rbind(stb1,stb2)
# unique(stb$Berry)
# unique(umnstb$type)
# unique(stb$measure)
# unique(stb$other)


# head(stb)
kable_styling(kable(head(stb, n=6)), font_size=12)

```

----

<font size ="5"> <p align="left">
&ensp; &ensp; &ensp;Berry column is a redundancy, delete it. Then do the same to separate Domain and Domain Category into left and right sides columns.
</p></font>
```{r}
stb <- select(stb, -Berry)

# Separate Domain
# unique(stb$Domain)
stb  <-  separate(stb,Domain, c("D_left", "D_right"), sep = ", ")
# unique(stb$D_left)
# unique(stb$D_right)

# Separate Domain Category
# unique(stb$`Domain Category`)

s1 <- str_detect(stb$`Domain Category`,"^CHEMICAL, INSECTICIDE: \\(CYFLUMETOFEN")

s2 <- which(s1 == FALSE)
stb2 <- stb[s2,]

s1 <- which(s1 == TRUE)
stb1 <- stb[s1,]
# unique(stb1$`Domain Category`)
stb1$`Domain Category` <- "CHEMICAL, INSECTICIDE: (CYFLUMETOFEN = 138831)"
# unique(stb1$`Domain Category`)

stb <- rbind(stb1,stb2)

stb <-  separate(stb,`Domain Category`, c("DC_left", "DC_right"), sep = ", ")
# unique(stb$DC_left)
# unique(stb$DC_right)

stb <- separate(stb, DC_left, c("DC_left_l", "DC_left_r"), sep = ": ")
stb <- separate(stb, DC_right, c("DC_right_l", "DC_right_r"), sep = ": ")

kable_styling(kable(head(stb, n=6)), font_size=10)
```

----
<font size ="5"> <p align="left">
&ensp; &ensp; &ensp;Check the D_left and DC_left_l column, they have the same value meaning, and the same to D_right and DC_right_l. 
</p></font>
<font size ="5"> <p align="left">
&ensp; &ensp; &ensp;Save D_left, D_right and and DC_left_r as a new variable: Chemical.
</p></font>


```{r}
# Organize data
# unique(paste(stb$D_left,stb$DC_left_l))
stb <- select(stb, -DC_left_l)

# unique(paste(stb$D_right,stb$DC_right_l))
stb <- select(stb, -DC_right_l)

# unique(paste(stb$D_left, stb$D_right))

stb <- mutate(stb, D_left = "CHEMICAL", D_left = "") 
stb[is.na(stb)] <- " "
stb <- mutate(stb, Chemical=paste(D_left, D_right)) 
stb <- select(stb, -c(D_left, D_right)) 
stb[is.na(stb)] <- " "
stb <- mutate(stb, Chemical = str_trim(paste(DC_left_r, Chemical)))
kable_styling(kable(head(stb, n=6)), font_size=12)
```


## Reorganize the data
<br>
<font size ="5"> <p align="left">
&ensp; &ensp; &ensp;Reorganize the strawberries data, select useful columns and rename the variables.
</p></font>
```{r}
stb <- select(stb, Year, State, type, measure, other, DC_right_r, Chemical, Value )
# unique(stb$type) # rename Production
# unique(stb$measure) # rename Measures
# unique(stb$other) # rename Avg
# unique(stb$DC_right_r) # rename materials
# unique(stb$Chemical)
# unique(stb$Value) # Values

stb <- rename(stb, Production = type, Measures = measure, Avg = other, Materials = DC_right_r, Values = Value)
stb <- select(stb, Year, State, Production, Measures, Avg, Materials, Chemical, Values)
# stb[is.na(stb)] <- " "
kable_styling(kable(head(stb, n=6)), font_size=12)
```

# Exploratory data analysis
<br>
<font size ="5"> <p align="left">
&ensp; &ensp; &ensp;1. Select data from strawberries
</p></font>
<font size ="5"> <p align="left">
&ensp; &ensp; &ensp;2. Chemical
</p></font>
<font size ="5"> <p align="left">
&ensp; &ensp; &ensp;3. Acres Planted
</p></font>

## Select data from strawberries
<font size ="4"> <p align="left">
&ensp; &ensp; &ensp;In order to do some exploratory analysis, select data from strawberries which does not equal to NA, D and Z. Transfer the value from character into numeric. 
</p></font>

```{r}
stb_d <- filter(stb,Values != "(NA)" & Values != "(D)" & Values != "(Z)")
stb_d$Values <- gsub(",", "",stb_d$Values)
stb_d$Values <- as.numeric(stb_d$Values)

stb_d$Year <- as.ordered(stb_d$Year)
stb_d$State <- as.factor(stb_d$State)
stb_d$Production <- as.factor(stb_d$Production)
stb_d$Measures <- as.factor(stb_d$Measures)
stb_d$Avg <- as.factor(stb_d$Avg)
stb_d$Materials <- as.factor(stb_d$Materials)
stb_d$Chemical <- as.factor(stb_d$Chemical)

kable_styling(kable(head(stb_d, n=6)), font_size=12)
```

<font size ="4"> <p align="left">
&ensp; &ensp; &ensp;&ensp;Also transfer all other variables into standard data type.
</p></font>

```{r}
str(stb_d)
```


## Chemical
<font size ="4"> <p align="left">
&ensp; &ensp; &ensp; &ensp;&ensp;&ensp;&ensp;&ensp;&ensp;&ensp;First look at the different chemical values.
</p></font>
```{r}
ay1 <- filter(stb_d, Measures == " MEASURED IN LB / ACRE / APPLICATION")

p1 <- ggplot(ay1, aes(x = Chemical, y = Values)) +
      geom_boxplot() +
      theme(axis.text.x = element_text(angle = 60, hjust = 1),
      axis.text = element_text(size = 11),
      axis.title = element_text(size = 13, face = "bold"))
# p1

ay2 <- filter(ay1, Chemical != "OTHER")

p2 <- ggplot(ay2, aes(x = Chemical, y = Values)) +
      geom_boxplot() +
      theme(axis.text.x = element_text(angle = 60, hjust = 1),
      axis.text = element_text(size = 11),
      axis.title = element_text(size = 13, face = "bold"))
# p2

grid.arrange(p1, p2, ncol=2)

```

----

<font size ="4"> <p align="left">
&ensp; &ensp; &ensp;&ensp;&ensp;&ensp;&ensp;&ensp;&ensp;&ensp;Also compare the chemical values by different states.
</p></font>

```{r}

p5 <- ggplot(ay2, aes(x = Chemical, y = Values, color = State)) +
      geom_boxplot() +
      theme(axis.text.x = element_text(angle = 60, hjust = 1),
      axis.text = element_text(size = 11),
      axis.title = element_text(size = 13, face = "bold"))
p5
```

## Acres Planted
<font size ="4"> <p align="left">
&ensp;&ensp;&ensp;&ensp;&ensp;&ensp;&ensp;&ensp;&ensp;&ensp;Compare acres planted values from different states.
</p></font>
```{r}
ay3 <- filter(stb_d, Production == " ACRES PLANTED")

p3 <- ggplot(ay3, aes(x = State, y = Values)) +
      geom_boxplot() +
      theme(axis.text.x = element_text(angle = 60, hjust = 1),
      axis.text = element_text(size = 11),
      axis.title = element_text(size = 13, face = "bold"))
# p3

ay4 <- filter(ay3, State != "CALIFORNIA")

p4 <- ggplot(ay4, aes(x = State, y = Values)) +
      geom_boxplot() +
      theme(axis.text.x = element_text(angle = 60, hjust = 1),
      axis.text = element_text(size = 11),
      axis.title = element_text(size = 13, face = "bold"))
# p4

grid.arrange(p3, p4, ncol=2)
```

# References

1. Edwin de Jonge, Mark van der Loo, An introduction to data cleaning with R.
2. Rstudio tutorial. https://shiny.rstudio.com/tutorial/written-tutorial/lesson6/
3. http://rstudio.github.io/shiny/tutorial/
4. Garrett Grolemund, Hadley Wickham, R for Data Science, https://r4ds.had.co.nz/
